\begingroup\fontsize{9}{11}\selectfont

\begin{ThreePartTable}
\begin{TableNotes}
\item \textit{Note: } 
\item This table compares vertical and horizontal distribution coefficients for different policy instruments across all countries. 
           Column '$\widehat{V}^{1} > 1$' displays the number of countries in which poorer households consume more carbon-intensively compared to richer households, under consideration of each policy.
           Column '$\hat{V}^{1} < 1$' displays the number of countries in which richer households consume more carbon-intensively compared to poorer households, under consideration of each policy.
           Column '$\hat{V}^{1} \uparrow$' displays the number of countries in which $\hat{V}^{1}$ increases in comparison to national climate policy, i.e., in which poorer households would consume more carbon-intensively compared to richer households and to the 'national climate policy'-scenario.
           Column '$\hat{V}^{1} \downarrow$' displays the number of countries in which $\hat{V}^{1}$ decreases in comparison to national climate policy, i.e., in which poorer households would consume less carbon-intensively compared to richer households and to the 'national climate policy'-scenario.
           Column '$\hat{H}^{1} > 1$' displays the number of countries in which carbon intensity is more heterogeneous among poorer households compared to richer households, under consideration of each policy.
           Column '$\hat{H}^{1} < 1$' displays the number of countries in which carbon intensity is more heterogeneous among richer households compared to poorer households, under consideration of each policy.
           Column '$\hat{H}^{1} \uparrow$' displays the number of countries in which $\hat{H}^{1}$ increases in comparison to national climate policy, i.e., in which heterogeneity among poorer households compared to richer households would increase in comparison to the 'national climate policy'-scenario.
           Column '$\hat{H}^{1} \downarrow$' displays the number of countries in which $\hat{H}^{1}$ decreases in comparison to national climate policy, i.e., in which heterogeneity among poorer households compared to richer households would decrease in comparison to the 'national climate policy'-scenario.
\end{TableNotes}
\begin{longtable}[t]{l|cc|cc|cc|ccl|cc|cc|cc|ccl|cc|cc|cc|ccl|cc|cc|cc|ccl|cc|cc|cc|ccl|cc|cc|cc|ccl|cc|cc|cc|ccl|cc|cc|cc|ccl|cc|cc|cc|cc}
\caption{\label{tab:A14}Comparing vertical and horizontal distribution coefficients for different policies}\\
\toprule
\multicolumn{1}{c}{Policy instrument} & \multicolumn{1}{c}{$\widehat{V}^{1} > 1$} & \multicolumn{1}{c}{$\widehat{V}^{1} < 1$} & \multicolumn{1}{c}{$\widehat{V}^{1} \uparrow $} & \multicolumn{1}{c}{$\widehat{V}^{1} \downarrow $} & \multicolumn{1}{c}{$\widehat{H}^{1} > 1$} & \multicolumn{1}{c}{$\widehat{H}^{1} < 1$} & \multicolumn{1}{c}{$\widehat{H}^{1} \uparrow $} & \multicolumn{1}{c}{$\widehat{H}^{1} \downarrow $} \\
\cmidrule(l{3pt}r{3pt}){1-1} \cmidrule(l{3pt}r{3pt}){2-2} \cmidrule(l{3pt}r{3pt}){3-3} \cmidrule(l{3pt}r{3pt}){4-4} \cmidrule(l{3pt}r{3pt}){5-5} \cmidrule(l{3pt}r{3pt}){6-6} \cmidrule(l{3pt}r{3pt}){7-7} \cmidrule(l{3pt}r{3pt}){8-8} \cmidrule(l{3pt}r{3pt}){9-9}
\endfirsthead
\caption[]{Comparing vertical and horizontal distribution coefficients for different policies \textit{(continued)}}\\
\toprule
\endhead

\endfoot
\bottomrule
\insertTableNotes
\endlastfoot
National climate policy & 44 & 44 &  &  & 60 & 28 &  & \\
International climate policy & 47 & 41 & 44 & 44 & 57 & 31 & 30 & 58\\
Transport sector policy & 29 & 59 & 10 & 78 & 50 & 38 & 28 & 60\\
Electricity sector policy & 62 & 26 & 74 & 14 & 65 & 23 & 63 & 25\\*
\end{longtable}
\end{ThreePartTable}
\endgroup{}
